---
title: "Rental Pricing Data Analysis"
output: html_notebook
---

# Rental Pricing Trend within Klang Valley
Dataset was taken from Kaggle, webscraped from Mudah.com by a public user. This project aims to provide visual insights on the trend of monthly rent with given category (provided in Mudah website) such as location, size, number of bedrooms, parking or furnished. 
This dataset is chosen as I am also actively looking for a new unit to move in Klang Valley. Therefore having this insights will hopefully help me compare my options and choices that I already had, to compare it within the actual market trend.

### Installing data packages and library

```{r}
install.packages("stringr")
library(stringr)
install.packages("ggplot2")
library(ggplot2)
install.packages("tidyverse")
library(tidyverse)
install.packages("dplyr")
library(dplyr)
library(readr)
```

```{r}
rental_df = read.csv("COURSERA DA/Rental Pricing.csv")
```

### Data Exploration

To describe the whole dataset (rows and columns), also checking on how many empty cells are within the dataset.

```{r}
colnames(rental_df)
```

```{r}
str(rental_df)
```

```{r}
head(rental_df)
```

```{r}
emptycell = colSums(is.na(rental_df))
print(emptycell)
```
Dataset is already good as it is. Empty cells in column 'parking' and 'completion_year' will be ignored as it will not be used as much in the upcoming analysis

### Data Preparation
Checking for numerical outliers in the dataset; size and monthly rent. Plotting graph with outliers will make our plots skewed.
```{r}
ggplot (data = rental_df) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  labs(title = "Plotted chart with presence of outliers")
```

```{r}
rentsummary = summary(rental_df$monthly_rent)
q1 = rentsummary["1st Qu."]
median = rentsummary['Median']
q3 = rentsummary["3rd Qu."]

lower_limit = q1 - 1.5 * IQR(rental_df$monthly_rent, na.rm = TRUE)
upper_limit = q3 + 1.5 * IQR(rental_df$monthly_rent, na.rm = TRUE)


print(lower_limit)
print(upper_limit)
```

```{r}
par(mfrow = c(1, 2))
boxplot(rental_df$monthly_rent, main = "Boxplot of Monthly Rent", ylab = "Monthly Rent (RM)", cex.main = 0.8)
boxplot(rental_df$monthly_rent, main = "Boxplot of Monthly Rent (outliers removed)", ylab = "Monthly Rent (RM)", ylim = c(lower_limit, upper_limit), cex.main = 0.8)
text(x = 1.25, y = 800, labels = "IQR Range is from \nRM 50 - RM 2850", cex = 0.5)
```

```{r}
sizesummary = summary(rental_df$size)
q1_size = sizesummary["1st Qu."]
median_size = sizesummary['Median']
q3_size = sizesummary["3rd Qu."]

size_lowerlimit = q1 - 1.5 * IQR(rental_df$size, na.rm = TRUE)
size_upperlimit = q3 + 1.5 * IQR(rental_df$size, na.rm = TRUE)


print(size_lowerlimit)
print(size_upperlimit)
```

```{r}
par(mfrow = c(1, 2))
boxplot(rental_df$size, main = "Boxplot of House Size", ylab = "Size (sqft)", cex.main = 0.8)
boxplot(rental_df$size, main = "Boxplot of House Size (outliers removed)", ylab = "Size (sqft)", ylim = c(size_lowerlimit,size_upperlimit), cex.main = 0.8)
text(x = 1.25, y = 1200, labels = "IQR Range is from \n659 sqft - 2241 sqft", cex = 0.5)
```

Creating new dataframe after removing outliers in our numerical column: monthly rent and size.
```{r}
filteredrental = rental_df %>%
  filter(monthly_rent <= 2850 & size <= 2241)

dim(filteredrental)
```
#### Data Visualization
```{r}
ggplot(data = filteredrental, aes(x ="", fill = region)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title = "Distribution of rental property availability in KL and Selangor", fill = "Region")
```

```{r}
location_count = (table(rental_df$district)) %>%
  as.data.frame() %>%
  arrange(desc(Freq))

location_count %>%
  filter(Freq >= 100) %>%
  ggplot(aes(x = Freq, y = Var1, size = Freq, color = Freq)) +
  geom_point() +
  labs(title = "Distribution of Rental Availability according to District", x = "Location", y = "Frequency") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 5))

```

```{r}
ggplot (data = filteredrental) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  facet_wrap(~region) + 
  labs(title = "Distribution of rental property according to Property Type")

```
Removing property type with too few data, to visually improving the plot.
```{r}
table(filteredrental$property_type)
```


```{r}
apt_rental = filteredrental %>%
  filter(property_type %in%  c("Apartment", "Condominium", "Service Residence", "Flat", "Studio", "Duplex"))

ggplot (data = apt_rental) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  facet_wrap(~region) +
  labs(title = "Distribution of rental property according to Property Type")
```
### Analysing rental dependency on other variable
```{r}
ggplot(filteredrental, aes(x = monthly_rent, fill = furnished)) +
  geom_density(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Rental dependency on Furnished Condition") +
  theme_minimal()

```

```{r}
ggplot(filteredrental, aes(x = monthly_rent, fill = region)) +
  geom_density(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Rental dependency on Region") +
  theme_minimal()
```

```{r}
ggplot(apt_rental, aes(x = monthly_rent, fill = property_type)) +
  geom_density(alpha = 0.4) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Rental dependency on Property Type") +
  theme_minimal()
```
Extracting "Near KTM/LRT" information into a new column, added onto dataframe.
```{r}
rental_publictransport = filteredrental %>%
  mutate(
    public_transport = str_extract(additional_facilities, "Near KTM/LRT"),
    public_transport = replace(public_transport, is.na(public_transport), "Not Near KTM/LRT"))
```

```{r}
ggplot(rental_publictransport, aes(x = monthly_rent, fill = public_transport)) +
  geom_density(alpha = 0.4) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Rental dependency on Distance to Public Transport") +
  theme_minimal()
```
These graphs may conclude that:
1) Location and Distance to Public Transport do not have major affect on rental distribution.
2) Furnished condition and Property type displays significant differences on rental distribution.

Further analysis to be done to gain insights on rental against furnished condition and property type.
```{r}
filteredrental %>%
  select(furnished, monthly_rent) %>%
  ggplot(aes(x = furnished, y = monthly_rent, fill = furnished)) +
  geom_boxplot() +
  labs(title = "Boxplot of Monthly Rent with Furnished Condition",
       subtitle = "Median rent for Fully Furnished = RM 1700 \nMedian rent for Not Furnished = RM 999 \nMedian rent for Partially Furnished = RM 1300") +
  theme(plot.margin = margin(b=30))
```

```{r}
median_data = filteredrental %>%
  group_by(furnished) %>%
  summarize(median_rent = median(monthly_rent))
print(median_data)
```

```{r}
apt_rental %>%
  select(property_type, monthly_rent) %>%
  ggplot(aes(x = property_type, y = monthly_rent, fill = property_type)) +
  geom_boxplot() +
  labs(title = "Boxplot of Monthly Rent with Property Type") +
  theme(plot.margin = margin(b=30))
```

```{r}
propertymedian_data = apt_rental %>%
  group_by(property_type) %>%
  summarize(propertymedian_rent = median(monthly_rent))
print(propertymedian_data)
```

### Summary
As a potential tenant, I'd always consider an affordable rent of below RM 1.5K as my budget, before deciding on a house. From this analysis, RM1.5K per month offers a lot of potential unit such as:
1) Apartment, flat, and condominium shows a median of monthyl rent close to RM 1.5k. However, service residence can be considered out of the budget.
2)Not furnished and partially furnished rental are still within RM 1.5k budget, however a fully furnished unit median is around RM1.7K.
3) Monthyl rental for units in Kl are slighly higher that units in Selangor.
4) Vicinity of public transport, whether it is near KTM/LRT or not, does not significantly affect monthly rental as much.
