---
title: "Data Analysis of Rental Pricing in Klang Valley"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: sentence
---

# Rental Pricing Trend within Klang Valley

Dataset was taken from Kaggle, webscraped from Mudah.com by a public user.
This project aims to provide visual insights on Klang Valley's rental trend with given category (provided in Mudah website) such as location, size, number of bedrooms, parking or furnished.
Chosen this project as I am also actively looking for a new unit to move in Klang Valley, having this insights will help me compare my options, evaluating them within the actual market trend.

### Installing data packages and library

```{r}
options(repos = "https://cloud.r-project.org")

```

```{r}
install.packages("stringr")
library(stringr)
install.packages("ggplot2")
library(ggplot2)
install.packages("tidyverse")
library(tidyverse)
install.packages("dplyr")
library(dplyr)
library(readr)
install.packages("wordcloud")
library(wordcloud)
install.packages("broom")
library(broom)
```

```{r}
rental_df = read.csv("Rental Pricing.csv")
```

### Data Exploration

**To describe the whole dataset (rows and columns), also checking on how many empty cells are within the dataset**

```{r}
colnames(rental_df)
```

```{r}
str(rental_df)
```

```{r}
head(rental_df)
```

```{r}
emptycell = colSums(is.na(rental_df))
print(emptycell)
```

Dataset is already good as it is.
Empty cells in column 'parking' and 'completion_year' will be ignored as it will not be used as much in the upcoming analysis.

### Data Preparation

**Checking for numerical outliers in the dataset** \### Checking for outliers and important variable column of interest; size and monthly rent.
Plotting graph with outliers will make our plots skewed.

```{r}
ggplot(data = rental_df) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  labs(title = "Plotted chart with presence of outliers")
```

```{r}
rentsummary = summary(rental_df$monthly_rent)
q1 = rentsummary["1st Qu."]
median = rentsummary['Median']
q3 = rentsummary["3rd Qu."]

lower_limit = q1 - 1.5 * IQR(rental_df$monthly_rent, na.rm = TRUE)
upper_limit = q3 + 1.5 * IQR(rental_df$monthly_rent, na.rm = TRUE)


print(lower_limit)
print(upper_limit)
```

```{r}
par(mfrow = c(1, 2))
boxplot(rental_df$monthly_rent, main = "Boxplot of Monthly Rent", ylab = "Monthly Rent (RM)", cex.main = 0.8)
boxplot(rental_df$monthly_rent, main = "Boxplot of Monthly Rent (outliers removed)", ylab = "Monthly Rent (RM)", ylim = c(lower_limit, upper_limit), cex.main = 0.8)
text(x = 1.25, y = 800, labels = "IQR Range is from \nRM 50 - RM 2850", cex = 0.5)
```

```{r}
sizesummary = summary(rental_df$size)
q1_size = sizesummary["1st Qu."]
median_size = sizesummary['Median']
q3_size = sizesummary["3rd Qu."]

size_lowerlimit = q1 - 1.5 * IQR(rental_df$size, na.rm = TRUE)
size_upperlimit = q3 + 1.5 * IQR(rental_df$size, na.rm = TRUE)


print(size_lowerlimit)
print(size_upperlimit)
```

```{r}
par(mfrow = c(1, 2))
boxplot(rental_df$size, main = "Boxplot of House Size", ylab = "Size (sqft)", cex.main = 0.8)
boxplot(rental_df$size, main = "Boxplot of House Size (outliers removed)", ylab = "Size (sqft)", ylim = c(size_lowerlimit,size_upperlimit), cex.main = 0.8)
text(x = 1.25, y = 1200, labels = "IQR Range is from \n659 sqft - 2241 sqft", cex = 0.5)
```

Creating new dataframe after removing outliers in our numerical column: monthly rent and size.

```{r}
filteredrental = rental_df %>%
  filter(monthly_rent <= 2850 & size <= 2241 & monthly_rent > 100 & size > 100)

dim(filteredrental)
```

```{r}
summary(filteredrental)
```

#### Data Visualization

**To easily illustrate trends in our dataset.**

```{r}
ggplot(data = filteredrental, aes(x ="", fill = region)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title = "Distribution of rental property availability in KL and Selangor", fill = "Region")
```

```{r}
facility_lists = strsplit(filteredrental$facilities, ",")

# Create a list to store the frequency counts
facility_counts = vector("list", length = length(facility_lists))

# Count the frequency of each variable in the 'facilities' column
for (i in seq_along(facility_lists)) {
  facility_counts[[i]] = table(facility_lists[[i]])
}

facility_characters = unlist(facility_lists)

# Count the frequency of each character
character_frequency = table(facility_characters)

facility_df = as.data.frame(character_frequency) %>%
  filter(!(facility_characters %in% c(11,10,6))) %>%
  arrange(desc(Freq))

wordcloud(words = facility_df$facility_characters, freq = facility_df$Freq,
          scale = c(2, 0.8), min.freq = 100, random.order = FALSE,
          rot.per = 0.2, colors = brewer.pal(6, "Dark2"),
          main = "Distribution of facilities mentioned")
```

Most listed rentals in Klang Valley are frequently associated with these facilities; security, playground, swimming pool and so on.

```{r}
location_count = (table(rental_df$district)) %>%
  as.data.frame() %>%
  arrange(desc(Freq))
```

```{r}
location_count %>%
  filter(Freq >= 100) %>%
  ggplot(aes(x = Freq, y = Var1, size = Freq, color = Freq)) +
  geom_point() +
  labs(title = "Distribution of Rental Availability according to District", x = "Location", y = "Frequency") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 5))

```

```{r}
ggplot (data = filteredrental) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  facet_wrap(~region) + 
  labs(title = "Distribution of rental property according to Property Type")

```

Removing property type listed with less than 50, to visually improving the plot and analysis.

```{r}
table(filteredrental$property_type)
```

```{r}
apt_rental = filteredrental %>%
  filter(property_type %in%  c("Apartment", "Condominium", "Service Residence", "Flat", "Studio", "Duplex"))

ggplot (data = apt_rental) +
  geom_point(mapping = aes(x = size, y = monthly_rent, color = property_type)) +
  facet_wrap(~region) +
  labs(title = "Distribution of rental property according to Property Type")
```

### Analysing rental dependency on variable of interest.

```{r}
ggplot(filteredrental, aes(x = monthly_rent, fill = furnished)) +
  geom_density(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Rental dependency on Furnished Condition") +
  theme_minimal()

```

```{r}
ggplot(filteredrental, aes(x = monthly_rent, fill = region)) +
  geom_density(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Rental dependency on Region") +
  theme_minimal()
```

```{r}
ggplot(apt_rental, aes(x = monthly_rent, fill = property_type)) +
  geom_density(alpha = 0.4) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Rental dependency on Property Type") +
  theme_minimal()
```

Extracting "Near KTM/LRT" information into a new column, added onto dataframe.

```{r}
rental_publictransport = filteredrental %>%
  mutate(
    public_transport = str_extract(additional_facilities, "Near KTM/LRT"),
    public_transport = replace(public_transport, is.na(public_transport), "Not Near KTM/LRT"))
```

```{r}
ggplot(rental_publictransport, aes(x = monthly_rent, fill = public_transport)) +
  geom_density(alpha = 0.4) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Rental dependency on Distance to Public Transport") +
  theme_minimal()
```

```{r}
filteredrental %>%
  filter(!is.na(furnished) & furnished != "") %>%
  select(furnished, monthly_rent) %>%
  ggplot(aes(x = furnished, y = monthly_rent, fill = furnished)) +
  geom_boxplot() +
  labs(title = "Boxplot of Monthly Rent with Furnished Condition",
       subtitle = "Median rent for Fully Furnished = RM 1700 \nMedian rent for Not Furnished = RM 1000 \nMedian rent for Partially Furnished = RM 1300") +
  theme(plot.margin = margin(b=30))
```

```{r}
median_data = filteredrental %>%
  filter(!is.na(furnished) & furnished != "") %>%
  group_by(furnished) %>%
  summarize(median_rent = median(monthly_rent))
print(median_data)
```

```{r}
apt_rental %>%
  select(property_type, monthly_rent) %>%
  ggplot(aes(x = property_type, y = monthly_rent, fill = property_type)) +
  geom_boxplot() +
  labs(title = "Boxplot of Monthly Rent with Property Type") +
  theme(plot.margin = margin(b=30))
```

```{r}
propertymedian_data = apt_rental %>%
  group_by(property_type) %>%
  summarize(propertymedian_rent = median(monthly_rent))
print(propertymedian_data)
```

**Insights gained from all the visuals generated** As a potential tenant, I'd always consider an affordable rent of below RM 1.5K as my budget, before deciding on a house.
From this analysis, RM1.5K per month offers a lot of potential unit such as: **1)** Apartment, flat, and condominium shows a median of monthly rent close to RM 1.5k.
However, service residence can be considered out of the budget.
**2)**Not furnished and partially furnished rental are still within RM 1.5k budget, however a fully furnished unit median is around RM1.7k **3)** Monthlyl rental for units in Kl are slighly higher that units in Selangor.
**4)** Vicinity of public transport, whether it is near KTM/LRT or not, does not significantly affect monthly rental as much.
**5)** Will proceed on using a machine learning model to explore more on variables correlation and feature importance.

### Model prediction using Random Forest

```{r}
ohe_df = filteredrental %>%
  filter(property_type %in%  c("Apartment", "Condominium", "Service Residence", "Flat", "Studio", "Duplex")) %>%
  mutate(
    public_transport = str_extract(additional_facilities, "Near KTM/LRT"),
    public_transport = replace(public_transport, is.na(public_transport), "Not Near KTM/LRT")) %>%
  filter(complete.cases(.))

install.packages("fastDummies")
library(fastDummies)
ohe_df = dummy_columns(ohe_df, select_column = c("property_type", "furnished", "region", "public_transport")) %>%
  select(-'ads_id')
```

Removing non-numeric column

```{r}
ohe_numerical = list()
for (col in names(ohe_df)) {
  if (is.numeric(ohe_df[[col]])) 
    ohe_numerical[[col]] = ohe_df[[col]]
}
ohe_num_df = data.frame(ohe_numerical)
```

```{r}
cor(ohe_num_df)
```

```{r}
correlation_table = cor(ohe_num_df)
cor_df = data.frame(variable = colnames(ohe_num_df), correlation = correlation_table)
heatmap(correlation_table, 
        col = colorRampPalette(c("blue", "white", "red"))(100),
        main = "Correlation Heatmap",
        xlab = "",
        ylab = "")

```

Train model

```{r}
install.packages("randomForest")
library(randomForest)
```

```{r}
set.seed(123)
train_index = sample(1:nrow(ohe_num_df), 0.8 * nrow(ohe_num_df))
train_data = ohe_num_df[train_index, ]
test_data = ohe_num_df[-train_index, ]

# Define the target variable and predictors
target_variable = "monthly_rent"
predictors = setdiff(names(ohe_num_df), target_variable)

# Train a Random Forest model
rental_model = randomForest(formula(paste(target_variable, "~", paste(predictors, collapse = "+"))),
                      data = train_data)

# Print the model summary
print(rental_model)
```

Evaluating feature importance on targeted variable of 'monthly rent'

```{r}
varImpPlot(rental_model)
```

```{r}
prediction = predict(rental_model, newdata = test_data)

actual = test_data$monthly_rent
ss_total = sum((actual - mean(actual))^2)
ss_residual = sum((actual - prediction)^2)
r_squared = 1 - (ss_residual / ss_total)

print(r_squared)
```

```{r}
# Create a data frame for plotting
prediction_df = data.frame(Actual = actual, Predicted = prediction)

# Create Scatter Plot
ggplot(prediction_df, aes(x = Predicted, y = Actual)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +  # Add diagonal line
  labs(title = "Predicted vs Actual Values", x = "Predicted", y = "Actual") +
  theme_minimal()
```

```{r}
head(prediction_df)
```

### Conclusion

Random Forest model r-squared value is at 0.6225412, which only represent 60% of accuracy instead of 100%.
Model may be needed to be revised, improved or undergo hyperparameter tuning (of which I'm still exploring).
As for now, this model has illustrated that top 3 variables that highly affected monthly rent in Klang Valley is the unit's size, it's furnished condition and the year it's build.
Personally, I always thought that public transport facilities (whether its within/outside vicinity) will highly affect rental price, but my analysis shows the contrast.
This also shows that Klang Valley is definitely a city of cars, and most its potential residents are car owner, hence why it doesnt affect rental price.
